# Reproducible Research: Peer Assessment 1
## [github.com/l0neJT](http://www.github.com/l0neJT)
## 14 June 2014

### Introduction
This assignment demonstrates the use of [R Markdown](http://rmarkdown.rstudio.com/) in conjunction with the [knitr package](http://cran.r-project.org/web/packages/knitr/index.html) to produce literate code. Forked from [rdpeng/RepData_PeerAssessment1](http://github.com/rdpeng/RepData_PeerAssessment1).

### Loading and preprocessing the data
1. Load packages [plyr](http://cran.r-project.org/web/packages/plyr/index.html) and [ggplot2](http://cran.r-project.org/web/packages/ggplot2/index.html)

    ```{r loadPackages}
    library(plyr)
    library(ggplot2)
    ```

2. Load data from file activity.csv

    ```{r loadDat, cache = TRUE}
    dat <- read.csv("./data/activity.csv")
    ```

3. Transform the 'date' column from factor into date
    
    ```{r transformDat, cache = TRUE}
    dat <- transform(dat, date = as.Date(date, format = "%Y-%m-%d"))
    str(dat)
    ```

4. Remove NAs for first analyses
    
    ```{r removeNAs, cache = TRUE}
    datNoNA <- dat[!is.na(dat$steps), ]
    ```

### What is mean total number of steps taken per day?
1. Summarise steps by date (excluding NAs)

    ```{r sumStepsDate}
    stepsDate <- ddply(datNoNA, "date", summarise, steps = sum(steps))
    ```

2. Chart steps by date as a bar plot with mean and median lines

    ```{r histStepsDate}
    histStepsDate <- function(data) {
        # Bar plot of steps by date without NAs
        p <- qplot(date, steps, data = data, geom = "bar", stat = "identity")
        
        # Horizontal line at mean
        pMean <- mean(data$steps)
        p <- p + geom_hline(yintercept = pMean, color = "red", size = 2)
        # Label mean line
        posX <- with(data, date[ceiling(length(date) / 2)])
        posY <- 20000
        label <- paste("Mean =", format(pMean, nsmall = 1))
        p <- p + annotate("text", x = posX, y = posY, label = label,
                          color = "red")
        
        # Horizontal line at median
        pMedian <- median(data$steps)
        p <- p + geom_hline(yintercept = pMedian, color = "blue", linetype = 2,
                            size = 1)
        # Label median line
        posY <- 19000
        label <- paste("Median =", format(pMedian, nsmall = 1))
        p <- p + annotate("text", x = posX, y = posY, label = label,
                          color = "blue")
        
        # Return plot
        p
    }
    histStepsDate(stepsDate) + labs(title = "Total Steps by Date (Remove NAs)")
    ```

### What is the average daily activity pattern?
1. Summarise steps by interval (excluding NAs)

    ```{r avgStepsInterval}
    stepsInterval <- ddply(datNoNA, "interval", summarise, steps = mean(steps))
    ```

2. Chart steps by interval as a line with max

    ```{r lineStepsInterval}
    lineStepsInterval <- function(data) {
        # Bar plot of steps by date without NAs
        p <- qplot(interval, steps, data = data, geom = "line")
        
        # Label max
        stepsMax <- max(data$steps)
        intvMax <- subset(data, steps == stepsMax, select = interval)[[1]]
        posX <- intvMax + 825L
        posY <- stepsMax
        label <- paste("Interval =", intvMax, "| Max Avg Steps = ",
                       format(stepsMax))
        p <- p + annotate("point", x = intvMax, y = stepsMax, color = "red",
                          size = 5)
        p <- p + annotate("text", x = posX, y = posY, color = "red",
                          label = label)
        
        # Return plot
        p
    }
    lineStepsInterval(stepsInterval) + labs(title =
                                            "Mean Steps by Interval (Remove NAs)")
    ```

### Imputing missing values
1. Replace NAs with mean for interval

    ``` {r replaceNAs, cache = TRUE}
    replaceNA <- function(sSteps, iInterval, data = stepsInterval) {
        ifelse(is.na(sSteps),
               subset(data, iInterval == interval, select = steps)[[1]],
               sSteps)
    }
    datReplaceNA <- ddply(dat, c("date", "interval"), transform,
                          steps = replaceNA(steps, interval))
    ```

2. Summarize steps by date and chart as a bar plot with mean and median lines

    ``` {r histReplaceNA}
    datTemp <- ddply(datReplaceNA, "date", summarise, steps = sum(steps))
    histStepsDate(datTemp) + labs(title = "Total Steps by Date (Replace NAs)")
    rm(datTemp)
    ```

3. Summarize steps by interval and chart as a line with max

    ```{r avgReplaceNA}
    datTemp <- ddply(datReplaceNA, "interval", summarise, steps = mean(steps))
    lineStepsInterval(datTemp) + labs(title =
                                      "Mean Steps by Interval (Replace NAs)")
    rm(datTemp)
    ```

### Are there differences in activity patterns between weekdays and weekends?
1. Add weekday character colum to data

    ``` {r addIsWeekday, cache = TRUE}
    weekDayOrEnd <- function(date) {
        ifelse(format(date, format = "%w") %in% 1:5,
               "weekday",
               "weekend")
    }
    datReplaceNA <- transform(datReplaceNA, weekday = weekDayOrEnd(date))
    ```

2. Summarise steps by interval categorized by weekday/weekend

    ```{r avgStepsSplit}
    stepsSplit <- ddply(datReplaceNA, c("interval", "weekday"), summarise,
                        steps = mean(steps))
    ```

3. Chart steps by interval as a line with facets by weekday/weekend

    ```{r lineStepsSplit}
    p <- qplot(interval, steps, data = stepsSplit, geom = c("line", "smooth"),
               facets = . ~ weekday)
    p + labs(title = "Mean Steps by Interval (Replace NAs)")
    rm(p)
    ```